beastscriptr: BEAUti for R

### 2017-10-09 article.tex
Figure 1

\usepackage{listings}
\usepackage{todonotes}

@@

@@

\begin{abstract}
  \textbf{Motivation:} BEAST2 is a popular Bayesian phylogenetics tool.
  The program BEAUti allows for a user-friendly way to create the
  input files that BEAST2 needs. This is convenient to get started, yet
  does not facilitate the scripted use of BEAST2. 
  beastscriptr is an R package that can create BEAST2 input files from an R
  function call.\\
  \textbf{Results:} beastscriptr has only parts of the functionality of BEAUti, yet
  can be extended easily. beastcriptr, on the other hand, does allow for specifying
  to use phylogenies of fixed crown age.\\
  \textbf{Availability:} The package is free and is available from the official R package archive at 
  http://cran.r-project.org/src/contrib/PACKAGES.html\#beastscriptr. 
  beastscriptr is licensed under the GNU General Public License.
  \textbf{Key-words:} computational biology, evolution, phylogeny, BEAST2, R
\end{abstract}

@@

BEAST2 \cite{bouckaert2014beast} is a popular Bayesian phylogenetics tool.
BEAST2 is bundled with the program called BEAUti \cite{drummond2012bayesian},
which  allows for a user-friendly way to create the
input files that BEAST2 needs. This is convenient to get started, yet
does not allow for the scripted use of BEAST2. beastscriptr is an R package 
that can create BEAST2 input files from an R function call.

@@

@@

@@

@@

\appendix

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Approach}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The goal of the R package beastscriptr is to create the 
same BEAST2 input files that BEAUti creates, but from one R function call 
instead.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methods}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

beastscriptr has a novel functionality not built into BEAUti yet:
it allows for using phylogenies of a fixed crown age.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\usepackage{listings}
\usepackage{todonotes}
\usepackage{verbatim}
\usepackage{tikz}
\usepackage{tkz-graph}
\usepackage{pgf}
\usetikzlibrary{arrows,automata}

% \newcommand*\beastscriptr{\verb;beastscriptr;}

@@

% \newcommand*\filename[1]{${#1}$}
% \newcommand*\myfilename[1]{\verb;{#1};}

@@

\begin{abstract}
  \textbf{1. }
  Here, I present a package, \verb;beastscriptr;, for the R programming language. \\
  \textbf{2. }
    \verb;beastscriptr; allows for scripted use of the BEAST2 phylogenetics tool, 
    by creating BEAST2 input files from an R function call. \\
  \textbf{3. }
    I describe \verb;beastscriptr; usage, the novel functionality it provides
    compared to BEAUti, and give some minimal examples. \\
  \textbf{4. }
    As \verb;beastscriptr; is free, libre, open-source and designed to be extended, 
    I conclude by describing the current development of the package \\
\end{abstract}

% Key-words: computational biology, evolution, phylogeny, BEAST2, R

@@

BEAST2 \cite{bouckaert2014beast} is a popular Bayesian phylogenetics tool.
BEAST2 is bundled with the program called BEAUti \cite{drummond2012bayesian},
which  allows for a user-friendly way to create the
input files that BEAST2 needs. BEAUti is convenient to get started, yet
does not allow for the scripted use of BEAST2. For creating BEAST2 input
files from a script, beastscriptr is the R package facilitating this:
it creates a BEAST2 input files from an R function call.

%\begin{figure}
%  \centering
%  \includegraphics[]{BeautiSiteModel.png}
%  \caption{BEAUti}
%  \label{figure:beauti}
%\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=6cm, semithick]   
  \tikzstyle{every state}=[]
  \node[state] (A) [rectangle] {
    \includegraphics[width=0.2\textwidth]{alignment.png}
    \includegraphics[width=0.2\textwidth]{thought_cloud.png}
  };   
  \node[state] (E) [below of=A, rectangle] {
    \includegraphics[height=0.1\textheight]{xml.png}
  };   
  \node[state] (F) [rectangle, below of=E] {\includegraphics[width=0.3\textwidth]{DensiTreeExample2.jpg}};
  \path (A) edge [anchor = east] node {\includegraphics[height=0.15\textheight]{beastscriptr_logo.png}} (E)
        (A) edge [anchor = west] node {\includegraphics[height=0.15\textheight]{BeautiSiteModel.png}} (E)
        (E) edge [] node {\includegraphics[width=0.2\textwidth]{Beast2LogoRat.png}} (F); 
  \end{tikzpicture}

  \caption{
    Workflow
  }
  \label{fig:beauti}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The goal of the R package \verb;beastscriptr; is to create the 
same BEAST2 input files that BEAUti creates, but from one R function call 
instead.

\verb;beastscriptr; has only parts of the functionality of BEAUti, yet
can be extended easily. beastcriptr, on the other hand, does allow for specifying
to use phylogenies of fixed crown age.

\verb;beastscriptr; has a novel functionality not built into BEAUti yet:
it allows for using phylogenies of a fixed crown age.

@@

As in BEAUti, to create a BEAST2 input file, 
one needs an alignment, model priors and MCMC parameters.
When the alignment is stored in the file alignment.fasta,
the desired tree prior is a constant-rate Birth-Death 
model \cite{nee1994reconstructed} and the MCMC chain needs to run 1M 
states, beastscriptr can create a file 
(called \verb;birth_death.xml;) from R 
as such:

\begin{lstlisting}[language=R]
beast_scriptr(
  input_fasta_filename = "alignment.fasta",
  mcmc_chainlength = 10000000,
  tree_prior = "birth_death",
  output_xml_filename = "birth_death.xml"
)
\end{lstlisting}

The created file \verb;birth_death.xml; can then be used by BEAST2.

Novel about beastscriptr is that it allows for specifying a fixed crown age:

\begin{lstlisting}[language=R]
beast_scriptr(
  input_fasta_filename = "alignment.fasta",
  mcmc_chainlength = 10000000,
  tree_prior = "birth_death",
  output_xml_filename = "birth_death_fixed_15.xml"
  fixed_crown_age = TRUE,
  initial_phylogeny = beastscriptr::fasta_to_phylo(
    input_fasta_filename, crown_age = 15)
)
\end{lstlisting}

In this example, a random phylogeny is created from the FASTA file, but
one could also use a more informed initial phylogeny with the desired crown age.

@@

The package is free and is available from the official R package archive at 
http://cran.r-project.org/src/contrib/PACKAGES.html\#beastscriptr. 
beastscriptr is licensed under the GNU General Public License.

@@
